Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping

### 2025-05-22 CaseStudies/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design of an integer-N frequency synthesizer, focusing on the theoretical and practical aspects of system dimensioning. The objective is to ensure precise frequency generation within a specified range, while meeting constraints related to frequency resolution and settling time. The design methodology involves the definition of key dynamic parameters, such as natural frequency, damping factor, and loop bandwidth, based on continuous-time control theory. A voltage-controlled oscillator (VCO) is configured to provide the required tuning range, accounting for process variations and voltage supply limitations. Additionally, a second-order passive loop filter is designed to ensure stability and appropriate dynamic response. The overall approach demonstrates a systematic framework for the design of robust and efficient frequency synthesizers.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

This report presents the design of an integer-N frequency synthesizer operating from `{python} format(f0)`Hz to `{python} format(f)`Hz. The synthesizer is required to achieve frequency settling within a time of $t_s=$ `{python} format(t_s)`s. Given that the channel spacing is `{python} format(f_ref)`Hz and an integer-N architecture is employed, the channel spacing is equal to the reference frequency. This example is based on the methodology and parameters provided in @rogers2006IntegratedCircuitDesign.

>>>>

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design of an integer-N frequency synthesizer, focusing on the theoretical and practical aspects of system dimensioning. The objective is to ensure precise frequency generation within a specified range, while meeting constraints related to frequency resolution and settling time. The design methodology involves the definition of key dynamic parameters, such as natural frequency, damping factor, and loop bandwidth, based on continuous-time control theory. A voltage-controlled oscillator (VCO) is configured to provide the required tuning range, accounting for process variations and voltage supply limitations. Additionally, a second-order passive loop filter is designed to ensure stability and appropriate dynamic response. The overall approach demonstrates a systematic framework for the design of simple frequency synthesizers.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

This report presents the design of an integer-N frequency synthesizer operating from `{python} format(f0)`Hz to `{python} format(f, 4)`Hz. The synthesizer is required to achieve frequency settling within a time of $t_s=$ `{python} format(t_s)`s. Given that the channel spacing is `{python} format(f_ref)`Hz and an integer-N architecture is employed, the channel spacing is equal to the reference frequency. This example is based on the methodology and parameters provided in @rogers2006IntegratedCircuitDesign.

### 2025-05-22 CaseStudies/Example3_4_SimulationandEstimationofLoopSettlingTimes/article/article.qmd

---
title: Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping
subtitle: Case Study - Example 3.4 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report explores the linearity limits of Phase-Frequency Detectors (PFDs) in Phase-Locked Loops (PLLs) and the onset of cycle slipping, a nonlinear phenomenon that occurs when the phase error exceeds the detector's effective operating range. The analysis focuses on tristate PFDs commonly used with charge pumps, emphasizing their functional behavior and limitations. A practical methodology for estimating the linear range through simulation is presented, along with a discussion on how exceeding this range can lead to loss of phase tracking and degraded loop performance. The consequences of cycle slipping, including increased jitter and potential loss of lock, are also examined.
keywords: 
    - Fractional-N PLL
    - Linearity Limit
    - Cycle Slipping
    - Nonlinear Dynamics
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

A fractional-N PLL synthesizer was designed using the provided specifications and components. The loop filter parameters were calculated based on continuous-time modeling, and the loop performance was evaluated under both linear and nonlinear frequency step scenarios. Settling times were estimated analytically, and results indicate acceptable behavior under the expected conditions. Further validation through simulation is necessary to confirm theoretical predictions.

>>>>

---
title: Estimation of Fractional-N PLL Linearity Limits and Cycle Slipping
subtitle: Case Study - Example 3.4 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design and analysis of a fractional-N frequency synthesizer based on a charge pump phase-locked loop (CP-PLL) architecture. The system comprises a voltage-controlled oscillator (VCO), a charge pump, a reference oscillator, and a second-order loop filter. Key parameters such as the loop bandwidth, natural frequency, and damping factor are used to guide the filter design. The settling time is estimated for different frequency steps, and the system’s response is evaluated for both linear and nonlinear operating regions. The methodology follows a continuous-time approach.
keywords: 
    - Fractional-N PLL
    - Linearity Limit
    - Cycle Slipping
    - Nonlinear Dynamics
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

A fractional-N PLL synthesizer was designed using the provided specifications and components. The loop filter parameters were calculated based on continuous-time modeling, and the loop performance was evaluated under both linear and nonlinear frequency step scenarios. Settling times were estimated analytically. Further validation through simulation is necessary to confirm theoretical predictions.
